Visualising How Non-linear Dimension Reduction Warps Your Data

### 2024-03-25 paper.qmd
Added 5 Gaussian Cluster 2D models

plot_list3_gau <- plot_PHATE_2D(PHATE_data_gau) + 
  geom_point(size = 0.00001, colour = "#4daf4a") +
  theme_linedraw() +
  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  annotate(geom = 'text', label = 'c', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)

@@

trimesh_removed_tsne_gau <- trimesh_removed_tsne_gau +
  theme_linedraw() +
  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  annotate(geom = 'text', label = 'b', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)

@@

## Decide by looking at MSE plot
num_bins_x_tsne_gau <- 60
num_bins_y_tsne_gau <- 79
hex_size_tsne_gau <- 0.01
### non-empty 229

hb_obj_tsne_gau <- hex_binning(data = umap_gau_scaled, x = "UMAP1", 
                      y = "UMAP2", num_bins_x = num_bins_x_tsne_gau, 
                      num_bins_y = num_bins_y_tsne_gau, x_start = NA, y_start = NA, 
                      buffer_x = NA, buffer_y = NA, hex_size = hex_size_tsne_gau, col_start = "UMAP")

@@

model_object_tsne_gau <- fit_highd_model( training_data = training_data_gau, 
                                 nldr_df_with_id = umap_gau_scaled, 
                                 x = "UMAP1", y = "UMAP2", 
                                 num_bins_x = num_bins_x_tsne_gau, 
                                 num_bins_y = num_bins_y_tsne_gau, 
                                 x_start = NA, y_start = NA, 
                                 buffer_x = NA, buffer_y = NA, 
                                 hex_size = hex_size_tsne_gau,
                                 is_rm_lwd_hex = FALSE, 
                                 benchmark_to_rm_lwd_hex = NA, 
                                 col_start_2d = "UMAP", 
                                 col_start_highd = "x")

trimesh_removed_tsne_gau <- trimesh_removed_tsne_gau +
  theme_linedraw() +
  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  annotate(geom = 'text', label = 'b', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)

# show_langevitour(df = df_all_tsne_gau, df_b = df_bin_tsne_gau, df_b_with_center_data = df_bin_centroids_tsne_gau, col_start = "x", distance_df = distance_tsne_gau, distance_col = "distance", benchmark_value = benchmark_tsne_gau)

  # find_non_empty_bins(data = umap_gau_scaled, x = "UMAP1", y = "UMAP2",
  #                   non_empty_bins = 198, x_start = NA, y_start = NA,
  #                   buffer_x = NA, buffer_y = NA, hex_size = 0.03,
  #                   col_start = "UMAP")

# hex_size_vec <- seq(0.02, 2, by = 0.01)
# 
# for (i in 1:length(hex_size_vec)) {
# 
#   find_non_empty_bins(data = umap_gau_scaled, x = "UMAP1", y = "UMAP2",
#                     non_empty_bins = 198, x_start = NA, y_start = NA,
#                     buffer_x = NA, buffer_y = NA, hex_size = hex_size_vec[i],
#                     col_start = "UMAP")
# 
# 
# }

@@

phate_gau_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = tibble::as_tibble(PHATE_data_gau), x = "PHATE1", y = "PHATE2"))) |>
  dplyr::rename(c("PHATE1" = "scaled_PHATE1",
                  "PHATE2" = "scaled_PHATE2")) |>
  dplyr::mutate(ID = 1:NROW(PHATE_data_gau))

## Decide by looking at MSE plot
num_bins_x_tsne_gau <- 60
num_bins_y_tsne_gau <- 156
hex_size_tsne_gau <- 0.01
### non-empty 179

hb_obj_tsne_gau <- hex_binning(data = phate_gau_scaled, x = "PHATE1", 
                      y = "PHATE2", num_bins_x = num_bins_x_tsne_gau, 
                      num_bins_y = num_bins_y_tsne_gau, x_start = NA, y_start = NA, 
                      buffer_x = NA, buffer_y = NA, hex_size = hex_size_tsne_gau, col_start = "PHATE")

## To obtain the standardise counts within hexbins
counts_df <- as.data.frame(do.call(cbind, hb_obj_tsne_gau$std_cts))
df_bin_centroids <- extract_hexbin_centroids(centroids_df = all_centroids_df, 
                                             counts_df = counts_df)
# ggplot(data = hex_grid, aes(x = x, y = y)) +
#   geom_polygon(fill = "white", color = "black", aes(group = hex_poly_id)) +
#   geom_point(data = all_centroids_df, aes(x = c_x, y = c_y), color = "red", size = 0.1) +
#   coord_fixed()

@@

model_object_tsne_gau <- fit_highd_model( training_data = training_data_gau, 
                                 nldr_df_with_id = phate_gau_scaled, 
                                 x = "PHATE1", y = "PHATE2", 
                                 num_bins_x = num_bins_x_tsne_gau, 
                                 num_bins_y = num_bins_y_tsne_gau, 
                                 x_start = NA, y_start = NA, 
                                 buffer_x = NA, buffer_y = NA, 
                                 hex_size = hex_size_tsne_gau,
                                 is_rm_lwd_hex = FALSE, 
                                 benchmark_to_rm_lwd_hex = NA, 
                                 col_start_2d = "PHATE", 
                                 col_start_highd = "x")

trimesh_removed_tsne_gau <- trimesh_removed_tsne_gau +
  theme_linedraw() +
  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  annotate(geom = 'text', label = 'b', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)

# show_langevitour(df = df_all_tsne_gau, df_b = df_bin_tsne_gau, df_b_with_center_data = df_bin_centroids_tsne_gau, col_start = "x", distance_df = distance_tsne_gau, distance_col = "distance", benchmark_value = benchmark_tsne_gau)

@@

## Decide by looking at MSE plot
num_bins_x_tsne_gau <- 46
num_bins_y_tsne_gau <- 61
hex_size_tsne_gau <- 0.013
### non-empty 179

hb_obj_tsne_gau <- hex_binning(data = trimap_gau_scaled, x = "TriMAP1", 
                      y = "TriMAP2", num_bins_x = num_bins_x_tsne_gau, 
                      num_bins_y = num_bins_y_tsne_gau, x_start = NA, y_start = NA, 
                      buffer_x = NA, buffer_y = NA, hex_size = hex_size_tsne_gau, col_start = "TriMAP")

@@

model_object_tsne_gau <- fit_highd_model( training_data = training_data_gau, 
                                 nldr_df_with_id = trimap_gau_scaled, 
                                 x = "TriMAP1", y = "TriMAP2", 
                                 num_bins_x = num_bins_x_tsne_gau, 
                                 num_bins_y = num_bins_y_tsne_gau, 
                                 x_start = NA, y_start = NA, 
                                 buffer_x = NA, buffer_y = NA, 
                                 hex_size = hex_size_tsne_gau,
                                 is_rm_lwd_hex = FALSE, 
                                 benchmark_to_rm_lwd_hex = NA, 
                                 col_start_2d = "TriMAP", 
                                 col_start_highd = "x")

trimesh_removed_tsne_gau <- trimesh_removed_tsne_gau +
  theme_linedraw() +
  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  annotate(geom = 'text', label = 'b', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)

# show_langevitour(df = df_all_tsne_gau, df_b = df_bin_tsne_gau, df_b_with_center_data = df_bin_centroids_tsne_gau, col_start = "x", distance_df = distance_tsne_gau, distance_col = "distance", benchmark_value = benchmark_tsne_gau)

@@

## Decide by looking at MSE plot
num_bins_x_tsne_gau <- 31
num_bins_y_tsne_gau <- 40
hex_size_tsne_gau <- 0.02
### non-empty 179

hb_obj_tsne_gau <- hex_binning(data = pacmap_gau_scaled, x = "PaCMAP1", 
                      y = "PaCMAP2", num_bins_x = num_bins_x_tsne_gau, 
                      num_bins_y = num_bins_y_tsne_gau, x_start = NA, y_start = NA, 
                      buffer_x = NA, buffer_y = NA, hex_size = hex_size_tsne_gau, col_start = "PaCMAP")

@@

model_object_tsne_gau <- fit_highd_model( training_data = training_data_gau, 
                                 nldr_df_with_id = pacmap_gau_scaled, 
                                 x = "PaCMAP1", y = "PaCMAP2", 
                                 num_bins_x = num_bins_x_tsne_gau, 
                                 num_bins_y = num_bins_y_tsne_gau, 
                                 x_start = NA, y_start = NA, 
                                 buffer_x = NA, buffer_y = NA, 
                                 hex_size = hex_size_tsne_gau,
                                 is_rm_lwd_hex = FALSE, 
                                 benchmark_to_rm_lwd_hex = NA, 
                                 col_start_2d = "PaCMAP", 
                                 col_start_highd = "x")

trimesh_removed_tsne_gau <- trimesh_removed_tsne_gau +
  theme_linedraw() +
  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  annotate(geom = 'text', label = 'b', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)

# show_langevitour(df = df_all_tsne_gau, df_b = df_bin_tsne_gau, df_b_with_center_data = df_bin_centroids_tsne_gau, col_start = "x", distance_df = distance_tsne_gau, distance_col = "distance", benchmark_value = benchmark_tsne_gau)

@@

<!---
```{r}
#| warning: false
#| echo: false
#| eval: false

### tSNE

# num_bins_tsne_gau <- calculate_effective_x_bins(.data = tSNE_data_gau, x = tSNE1,
#                                                 cell_area = 1) #56
num_bins_tsne_gau <- 16

shape_val_tsne_gau <- calculate_effective_shape_value(.data = tSNE_data_gau,
                                                      x = tSNE1, y = tSNE2)
## To extract bin centroids
hexbin_data_object_tsne_gau <- extract_hexbin_centroids(nldr_df = tSNE_data_gau, num_bins = num_bins_tsne_gau, shape_val = shape_val_tsne_gau, x = tSNE1, y = tSNE2)

df_bin_centroids_tsne_gau <- hexbin_data_object_tsne_gau$hexdf_data

num_non_empty_bins_gau <- df_bin_centroids_tsne_gau$hexID |> length()

tSNE_data_with_hb_id_gau <- tSNE_data_gau |>
  dplyr::mutate(hb_id = hexbin_data_object_tsne_gau$hb_data@cID)

## To generate a data set with high-D and 2D training data
df_all_tsne_gau <- dplyr::bind_cols(training_data_1 |> dplyr::select(-ID), tSNE_data_with_hb_id_gau)

## Averaged on high-D
df_bin_tsne_gau <- avg_highD_data(.data = df_all_tsne_gau)

## Triangulate bin centroids
tr1_object_tsne_gau <- triangulate_bin_centroids(df_bin_centroids_tsne_gau, x, y)
tr_from_to_df_tsne_gau <- generate_edge_info(triangular_object = tr1_object_tsne_gau)

## Compute 2D distances
distance_tsne_gau <- cal_2d_dist(.data = tr_from_to_df_tsne_gau)

## To find the benchmark value
benchmark_tsne_gau <- find_benchmark_value(.data = distance_tsne_gau, distance_col = distance)

trimesh_removed_tsne_gau <- remove_long_edges(.data = distance_tsne_gau, benchmark_value = benchmark_tsne_gau,
                                                  triangular_object = tr1_object_tsne_gau, distance_col = distance)

trimesh_removed_tsne_gau <- trimesh_removed_tsne_gau +
  # xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
  # theme(axis.text = element_text(size = 5),
  #       axis.title = element_text(size = 7)) +
  geom_point(colour = "#33a02c", size = 0.05) +
  theme_linedraw() +
  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)

# tour_tsne_gau <- show_langevitour(df_all_tsne_gau, df_bin_tsne_gau, df_bin_centroids_tsne_gau, benchmark_value = benchmark_tsne_gau, distance = distance_tsne_gau, distance_col = distance)

### UMAP

# num_bins_umap_gau <- calculate_effective_x_bins(.data = UMAP_data_gau, x = UMAP1,
#                                                 cell_area = 1) #41

>>>>

plot_list3_gau <- plot_PHATE_2D(PHATE_data_gau) + 
  geom_point(size = 0.00001, colour = "#4daf4a") +
  theme_linedraw() +
  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  annotate(geom = 'text', label = 'c', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)

@@

trimesh_removed_tsne_gau <- trimesh_removed_tsne_gau +
  theme_linedraw() +
  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)

@@

## Decide by looking at MSE plot
num_bins_x_umap_gau <- 60
num_bins_y_umap_gau <- 79
hex_size_umap_gau <- 0.01
### non-empty 229

hb_obj_umap_gau <- hex_binning(data = umap_gau_scaled, x = "UMAP1", 
                               y = "UMAP2", num_bins_x = num_bins_x_umap_gau, 
                               num_bins_y = num_bins_y_umap_gau, x_start = NA, y_start = NA, 
                               buffer_x = NA, buffer_y = NA, hex_size = hex_size_umap_gau, col_start = "UMAP")

## Data set with all possible centroids in the hexagonal grid
all_centroids_df <- as.data.frame(do.call(cbind, hb_obj_umap_gau$centroids))

## Generate all coordinates of hexagons
hex_grid <- as.data.frame(do.call(cbind, hb_obj_umap_gau$hex_poly))

## To obtain the standardise counts within hexbins
counts_df <- as.data.frame(do.call(cbind, hb_obj_umap_gau$std_cts))
df_bin_centroids <- extract_hexbin_centroids(centroids_df = all_centroids_df, 
                                             counts_df = counts_df)
# ggplot(data = hex_grid, aes(x = x, y = y)) +
#   geom_polygon(fill = "white", color = "black", aes(group = hex_poly_id)) +
#   geom_point(data = all_centroids_df, aes(x = c_x, y = c_y), color = "red") +
#   coord_fixed()

@@

umap_data_with_hb_id_gau <- as.data.frame(do.call(cbind, hb_obj_umap_gau$data_hb_id))

model_object_umap_gau <- fit_highd_model( training_data = training_data_gau, 
                                          nldr_df_with_id = umap_gau_scaled, 
                                          x = "UMAP1", y = "UMAP2", 
                                          num_bins_x = num_bins_x_umap_gau, 
                                          num_bins_y = num_bins_y_umap_gau, 
                                          x_start = NA, y_start = NA, 
                                          buffer_x = NA, buffer_y = NA, 
                                          hex_size = hex_size_umap_gau,
                                          is_rm_lwd_hex = FALSE, 
                                          benchmark_to_rm_lwd_hex = NA, 
                                          col_start_2d = "UMAP", 
                                          col_start_highd = "x")

df_bin_centroids_umap_gau <- model_object_umap_gau$df_bin_centroids
df_bin_umap_gau <- model_object_umap_gau$df_bin

## Triangulate bin centroids
tr1_object_umap_gau <- tri_bin_centroids(df_bin_centroids_umap_gau, x = "c_x", y = "c_y")
tr_from_to_df_umap_gau <- gen_edges(tri_object = tr1_object_umap_gau)

## Compute 2D distances
distance_umap_gau <- cal_2d_dist(tr_coord_df = tr_from_to_df_umap_gau, 
                                 start_x = "x_from", start_y = "y_from", 
                                 end_x = "x_to", end_y = "y_to", 
                                 select_vars = c("from", "to", "distance"))

# distance_plot <- plot_dist(distance_umap_gau) +
#   ylab(expression(d^{(2)})) +
#   theme(axis.text = element_text(size = 5),
#         axis.title = element_text(size = 12))
# 
# distance_plot

## To find the benchmark value
benchmark_umap_gau <- find_lg_benchmark(distance_edges = distance_umap_gau, distance_col = "distance")

# ggplot() +
# geom_trimesh(data = df_bin_centroids_umap_gau, mapping = aes(x = c_x, y = c_y))

trimesh_removed_umap_gau <- vis_rmlg_mesh(distance_edges = distance_umap_gau, benchmark_value = benchmark_umap_gau, tr_coord_df = tr_from_to_df_umap_gau, distance_col = "distance")

trimesh_removed_umap_gau <- trimesh_removed_umap_gau +
  theme_linedraw() +
  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  annotate(geom = 'text', label = 'b', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)

df_all_umap_gau <- dplyr::bind_cols(training_data_gau |> dplyr::select(-ID), umap_data_with_hb_id_gau)

# show_langevitour(df = df_all_umap_gau, df_b = df_bin_umap_gau, df_b_with_center_data = df_bin_centroids_umap_gau, col_start = "x", distance_df = distance_umap_gau, distance_col = "distance", benchmark_value = benchmark_umap_gau)
```

@@

phate_gau_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = tibble::as_tibble(PHATE_data_gau), x = "PHATE1", y = "PHATE2", hex_ratio = 50/sqrt(3)))) |>
  dplyr::rename(c("PHATE1" = "scaled_PHATE1",
                  "PHATE2" = "scaled_PHATE2")) |>
  dplyr::mutate(ID = 1:NROW(PHATE_data_gau))

## Decide by looking at MSE plot
num_bins_x_phate_gau <- 60
num_bins_y_phate_gau <- 1927
hex_size_phate_gau <- 0.01
### non-empty 187

hb_obj_phate_gau <- hex_binning(data = phate_gau_scaled, x = "PHATE1", 
                               y = "PHATE2", num_bins_x = num_bins_x_phate_gau, 
                               num_bins_y = num_bins_y_phate_gau, x_start = NA, y_start = NA, 
                               buffer_x = NA, buffer_y = NA, hex_size = hex_size_phate_gau, col_start = "PHATE")

## Data set with all possible centroids in the hexagonal grid
all_centroids_df <- as.data.frame(do.call(cbind, hb_obj_phate_gau$centroids))

## Generate all coordinates of hexagons
hex_grid <- as.data.frame(do.call(cbind, hb_obj_phate_gau$hex_poly))

## To obtain the standardise counts within hexbins
counts_df <- as.data.frame(do.call(cbind, hb_obj_phate_gau$std_cts))
df_bin_centroids <- extract_hexbin_centroids(centroids_df = all_centroids_df, 
                                             counts_df = counts_df)
# ggplot(data = hex_grid, aes(x = x, y = y)) +
#   geom_polygon(fill = "white", color = "black", aes(group = hex_poly_id)) +
#   geom_point(data = all_centroids_df, aes(x = c_x, y = c_y), color = "red", size = 0.1) +
#   coord_fixed()

@@

phate_data_with_hb_id_gau <- as.data.frame(do.call(cbind, hb_obj_phate_gau$data_hb_id))

model_object_phate_gau <- fit_highd_model( training_data = training_data_gau, 
                                          nldr_df_with_id = phate_gau_scaled, 
                                          x = "PHATE1", y = "PHATE2", 
                                          num_bins_x = num_bins_x_phate_gau, 
                                          num_bins_y = num_bins_y_phate_gau, 
                                          x_start = NA, y_start = NA, 
                                          buffer_x = NA, buffer_y = NA, 
                                          hex_size = hex_size_phate_gau,
                                          is_rm_lwd_hex = FALSE, 
                                          benchmark_to_rm_lwd_hex = NA, 
                                          col_start_2d = "PHATE", 
                                          col_start_highd = "x")

df_bin_centroids_phate_gau <- model_object_phate_gau$df_bin_centroids
df_bin_phate_gau <- model_object_phate_gau$df_bin

## Triangulate bin centroids
tr1_object_phate_gau <- tri_bin_centroids(df_bin_centroids_phate_gau, x = "c_x", y = "c_y")
tr_from_to_df_phate_gau <- gen_edges(tri_object = tr1_object_phate_gau)

## Compute 2D distances
distance_phate_gau <- cal_2d_dist(tr_coord_df = tr_from_to_df_phate_gau, 
                                 start_x = "x_from", start_y = "y_from", 
                                 end_x = "x_to", end_y = "y_to", 
                                 select_vars = c("from", "to", "distance"))

# distance_plot <- plot_dist(distance_phate_gau) +
#   ylab(expression(d^{(2)})) +
#   theme(axis.text = element_text(size = 5),
#         axis.title = element_text(size = 12))
# 
# distance_plot

## To find the benchmark value
benchmark_phate_gau <- find_lg_benchmark(distance_edges = distance_phate_gau, distance_col = "distance")

# ggplot() +
# geom_trimesh(data = df_bin_centroids_phate_gau, mapping = aes(x = c_x, y = c_y))

trimesh_removed_phate_gau <- vis_rmlg_mesh(distance_edges = distance_phate_gau, benchmark_value = benchmark_phate_gau, tr_coord_df = tr_from_to_df_phate_gau, distance_col = "distance")

trimesh_removed_phate_gau <- trimesh_removed_phate_gau +
  theme_linedraw() +
  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  annotate(geom = 'text', label = 'c', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)

df_all_phate_gau <- dplyr::bind_cols(training_data_gau |> dplyr::select(-ID), phate_data_with_hb_id_gau)

# show_langevitour(df = df_all_phate_gau, df_b = df_bin_phate_gau, df_b_with_center_data = df_bin_centroids_phate_gau, col_start = "x", distance_df = distance_phate_gau, distance_col = "distance", benchmark_value = benchmark_phate_gau)

@@

## Decide by looking at MSE plot
num_bins_x_trimap_gau <- 46
num_bins_y_trimap_gau <- 61
hex_size_trimap_gau <- 0.013
### non-empty 173

hb_obj_trimap_gau <- hex_binning(data = trimap_gau_scaled, x = "TriMAP1", 
                               y = "TriMAP2", num_bins_x = num_bins_x_trimap_gau, 
                               num_bins_y = num_bins_y_trimap_gau, x_start = NA, y_start = NA, 
                               buffer_x = NA, buffer_y = NA, hex_size = hex_size_trimap_gau, col_start = "TriMAP")

## Data set with all possible centroids in the hexagonal grid
all_centroids_df <- as.data.frame(do.call(cbind, hb_obj_trimap_gau$centroids))

## Generate all coordinates of hexagons
hex_grid <- as.data.frame(do.call(cbind, hb_obj_trimap_gau$hex_poly))

## To obtain the standardise counts within hexbins
counts_df <- as.data.frame(do.call(cbind, hb_obj_trimap_gau$std_cts))
df_bin_centroids <- extract_hexbin_centroids(centroids_df = all_centroids_df, 
                                             counts_df = counts_df)
# ggplot(data = hex_grid, aes(x = x, y = y)) +
#   geom_polygon(fill = "white", color = "black", aes(group = hex_poly_id)) +
#   geom_point(data = all_centroids_df, aes(x = c_x, y = c_y), color = "red") +
#   coord_fixed()

@@

trimap_data_with_hb_id_gau <- as.data.frame(do.call(cbind, hb_obj_trimap_gau$data_hb_id))

model_object_trimap_gau <- fit_highd_model( training_data = training_data_gau, 
                                          nldr_df_with_id = trimap_gau_scaled, 
                                          x = "TriMAP1", y = "TriMAP2", 
                                          num_bins_x = num_bins_x_trimap_gau, 
                                          num_bins_y = num_bins_y_trimap_gau, 
                                          x_start = NA, y_start = NA, 
                                          buffer_x = NA, buffer_y = NA, 
                                          hex_size = hex_size_trimap_gau,
                                          is_rm_lwd_hex = FALSE, 
                                          benchmark_to_rm_lwd_hex = NA, 
                                          col_start_2d = "TriMAP", 
                                          col_start_highd = "x")

df_bin_centroids_trimap_gau <- model_object_trimap_gau$df_bin_centroids
df_bin_trimap_gau <- model_object_trimap_gau$df_bin

## Triangulate bin centroids
tr1_object_trimap_gau <- tri_bin_centroids(df_bin_centroids_trimap_gau, x = "c_x", y = "c_y")
tr_from_to_df_trimap_gau <- gen_edges(tri_object = tr1_object_trimap_gau)

## Compute 2D distances
distance_trimap_gau <- cal_2d_dist(tr_coord_df = tr_from_to_df_trimap_gau, 
                                 start_x = "x_from", start_y = "y_from", 
                                 end_x = "x_to", end_y = "y_to", 
                                 select_vars = c("from", "to", "distance"))

# distance_plot <- plot_dist(distance_trimap_gau) +
#   ylab(expression(d^{(2)})) +
#   theme(axis.text = element_text(size = 5),
#         axis.title = element_text(size = 12))
# 
# distance_plot

## To find the benchmark value
benchmark_trimap_gau <- find_lg_benchmark(distance_edges = distance_trimap_gau, distance_col = "distance")
benchmark_trimap_gau <- 0.2

# ggplot() +
# geom_trimesh(data = df_bin_centroids_trimap_gau, mapping = aes(x = c_x, y = c_y))

trimesh_removed_trimap_gau <- vis_rmlg_mesh(distance_edges = distance_trimap_gau, benchmark_value = benchmark_trimap_gau, tr_coord_df = tr_from_to_df_trimap_gau, distance_col = "distance")

trimesh_removed_trimap_gau <- trimesh_removed_trimap_gau +
  theme_linedraw() +
  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  annotate(geom = 'text', label = 'd', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)

df_all_trimap_gau <- dplyr::bind_cols(training_data_gau |> dplyr::select(-ID), trimap_data_with_hb_id_gau)

# show_langevitour(df = df_all_trimap_gau, df_b = df_bin_trimap_gau, df_b_with_center_data = df_bin_centroids_trimap_gau, col_start = "x", distance_df = distance_trimap_gau, distance_col = "distance", benchmark_value = benchmark_trimap_gau)

@@

## Decide by looking at MSE plot
num_bins_x_pacmap_gau <- 31
num_bins_y_pacmap_gau <- 40
hex_size_pacmap_gau <- 0.02
### non-empty 179

hb_obj_pacmap_gau <- hex_binning(data = pacmap_gau_scaled, x = "PaCMAP1", 
                               y = "PaCMAP2", num_bins_x = num_bins_x_pacmap_gau, 
                               num_bins_y = num_bins_y_pacmap_gau, x_start = NA, y_start = NA, 
                               buffer_x = NA, buffer_y = NA, hex_size = hex_size_pacmap_gau, col_start = "PaCMAP")

## Data set with all possible centroids in the hexagonal grid
all_centroids_df <- as.data.frame(do.call(cbind, hb_obj_pacmap_gau$centroids))

## Generate all coordinates of hexagons
hex_grid <- as.data.frame(do.call(cbind, hb_obj_pacmap_gau$hex_poly))

## To obtain the standardise counts within hexbins
counts_df <- as.data.frame(do.call(cbind, hb_obj_pacmap_gau$std_cts))
df_bin_centroids <- extract_hexbin_centroids(centroids_df = all_centroids_df, 
                                             counts_df = counts_df)
# ggplot(data = hex_grid, aes(x = x, y = y)) +
#   geom_polygon(fill = "white", color = "black", aes(group = hex_poly_id)) +
#   geom_point(data = all_centroids_df, aes(x = c_x, y = c_y), color = "red") +
#   coord_fixed()

@@

pacmap_data_with_hb_id_gau <- as.data.frame(do.call(cbind, hb_obj_pacmap_gau$data_hb_id))

model_object_pacmap_gau <- fit_highd_model( training_data = training_data_gau, 
                                          nldr_df_with_id = pacmap_gau_scaled, 
                                          x = "PaCMAP1", y = "PaCMAP2", 
                                          num_bins_x = num_bins_x_pacmap_gau, 
                                          num_bins_y = num_bins_y_pacmap_gau, 
                                          x_start = NA, y_start = NA, 
                                          buffer_x = NA, buffer_y = NA, 
                                          hex_size = hex_size_pacmap_gau,
                                          is_rm_lwd_hex = FALSE, 
                                          benchmark_to_rm_lwd_hex = NA, 
                                          col_start_2d = "PaCMAP", 
                                          col_start_highd = "x")

df_bin_centroids_pacmap_gau <- model_object_pacmap_gau$df_bin_centroids
df_bin_pacmap_gau <- model_object_pacmap_gau$df_bin

## Triangulate bin centroids
tr1_object_pacmap_gau <- tri_bin_centroids(df_bin_centroids_pacmap_gau, x = "c_x", y = "c_y")
tr_from_to_df_pacmap_gau <- gen_edges(tri_object = tr1_object_pacmap_gau)

## Compute 2D distances
distance_pacmap_gau <- cal_2d_dist(tr_coord_df = tr_from_to_df_pacmap_gau, 
                                 start_x = "x_from", start_y = "y_from", 
                                 end_x = "x_to", end_y = "y_to", 
                                 select_vars = c("from", "to", "distance"))

# distance_plot <- plot_dist(distance_pacmap_gau) +
#   ylab(expression(d^{(2)})) +
#   theme(axis.text = element_text(size = 5),
#         axis.title = element_text(size = 12))
# 
# distance_plot

## To find the benchmark value
benchmark_pacmap_gau <- find_lg_benchmark(distance_edges = distance_pacmap_gau, distance_col = "distance")

# ggplot() +
# geom_trimesh(data = df_bin_centroids_pacmap_gau, mapping = aes(x = c_x, y = c_y))

trimesh_removed_pacmap_gau <- vis_rmlg_mesh(distance_edges = distance_pacmap_gau, benchmark_value = benchmark_pacmap_gau, tr_coord_df = tr_from_to_df_pacmap_gau, distance_col = "distance")

trimesh_removed_pacmap_gau <- trimesh_removed_pacmap_gau +
  theme_linedraw() +
  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  annotate(geom = 'text', label = 'e', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)

df_all_pacmap_gau <- dplyr::bind_cols(training_data_gau |> dplyr::select(-ID), pacmap_data_with_hb_id_gau)

# show_langevitour(df = df_all_pacmap_gau, df_b = df_bin_pacmap_gau, df_b_with_center_data = df_bin_centroids_pacmap_gau, col_start = "x", distance_df = distance_pacmap_gau, distance_col = "distance", benchmark_value = benchmark_pacmap_gau)

@@

```{r}
#| warning: false
#| echo: false
#| label: fig-modelfiveGau
#| fig-cap: "Is there a best model to represent the original data in 2D space or are they all providing equivalent information?, (a) Model in the 2D space with tSNE (<https://youtu.be/x5VPB-wOLm4>), (b) Model in the 2D space with UMAP (<https://youtu.be/Cm1dW6iLyHQ>), (c) Model in the 2D space with PHATE (<https://youtu.be/HB0Y3ilz6sI>), (d) Model in the 2D space with TriMAP (<https://youtu.be/iSVkXIrOrSA>), and (e) Model in the 2D space with PaCMAP (<https://youtu.be/O2QBctsc4uM>)."
#| out-width: 100%
